ArKanjo: a tool for detecting function-level Code Duplication in the Linux Kernel

### 2025-06-18 paper/paper.tex
Adding footnote about Free Software, and other minor adjustments

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Luan Ícaro Pinto Arcanjo}
\IEEEauthorblockA{
Universidade de São Paulo, Brazil\\
luanicaro@usp.br}
\and
\IEEEauthorblockN{David Tadokoro}
\IEEEauthorblockA{
Universidade de São Paulo, Brazil\\
davidbtadokoro@usp.br}
\and
\IEEEauthorblockN{Paulo Meirelles}
\IEEEauthorblockA{
Universidade de São Paulo, Brazil\\
paulormm@ime.usp.br}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
The Linux kernel massive scale (+28 M LoC, +20 K contributors) presents unique 
maintenance challenges. Surprisingly, code duplication remains a persistent issue 
in the kernel codebase, which can hinder its evolution and patching. Academic 
approaches often focus on pairwise comparison of code artifacts, which are not directly applied 
to comprehensive codebase analyses. Other existing free software tools explored in 
practice frequently suffer from limited functionality, such as primitive textual 
matching, prove too narrow in scope, or fail to deliver effective results on complex, 
large-scale codebases. Existing solutions generally fail to address the Linux kernel 
specific needs: (1) scalability to handle its size, (2) actionable results for 
developers, and (3) integration with kernel development workflows.
%
This paper presents \textit{ArKanjo}, a novel command-line tool for Linux kernel maintenance 
designed to detect and analyze function-level duplications. Released under the MIT 
license, ArKanjo employs a two-stage architecture consisting of a Preprocessor and 
a Query Responder that separates computationally intensive analysis from efficient 
querying for duplications within large codebases. Pivotal advantages of ArKanjo over 
existing solutions include: (1) optimization for C codebases with kernel-specific 
patterns; (2) preprocessing that enables rapid queries without redundant analysis; 
and (3) prioritization of duplicates that impact maintainability, such as copied 
buggy logic. 
%
We evaluate ArKanjo against real-world duplication cases in recent 
kernel versions, demonstrating its effectiveness in identifying problematic clones 
that generic tools often overlook. By identifying well-defined, manageable 
duplication instances, ArKanjo effectively lowers the barrier for new contributors, 
a capability evidenced by its role in guiding students to make their first code 
improvements to the kernel. ArKanjo offers immediate value to kernel maintainers 
and serves as a replicable model for clone detection in other large-scale free 
software projects.

@@

To fill this gap, this research proposes a new approach embodied in \textit{ArKanjo}, a tool 
designed specifically to identify and facilitate the mitigation of function-level code 
duplications within the Linux kernel. To test the tool, we employ multiple analytical 
methods and ethnographic studies.

@@

\begin{figure}[!t]
\centering
\includegraphics[width=3.5in]{explorer_example}
\caption{Example of the \textit{Duplication Explorer} functionality of the ArKanjo tool.}
\label{fig:explorer_ex}
\end{figure}

\begin{figure}[!t]
\centering
\includegraphics[width=3.5in]{function_example}
\caption{Example of the \textit{Function Information} functionality of the ArKanjo tool.}
\label{fig:function_ex}
\end{figure}

\begin{figure}[!t]
\centering
\includegraphics[width=3.5in]{relatory_example}
\caption{Example of the \textit{Duplication Report} functionality of the ArKanjo tool.}
\label{fig:relatory_ex}
\end{figure}

@@

The Duplication Report functionality provides an overview of
duplicated code within the input codebase. It calculates the number of
duplicated lines per folder in the codebase and presents the information to the user in
a readable format. Figure~\ref{fig:relatory_ex} demonstrates an example of this functionality in use.

@@

Nevertheless, a discrepancy emerges when detecting non-duplicate pairs between
the two methods. In the BigCloneBench evaluation, we found no false positives (i.e., negative
samples inferred as duplications), whereas the empirical analysis revealed a higher percentage
of false positives. We propose two potential explanations for this discrepancy.

>>>>

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Luan Arcanjo}
\IEEEauthorblockA{
Universidade de São Paulo, Brazil\\
luanicaro@usp.br}
\and
\IEEEauthorblockN{David Tadokoro}
\IEEEauthorblockA{
Universidade de São Paulo, Brazil\\
davidbtadokoro@usp.br}
\and
\IEEEauthorblockN{Paulo Meirelles}
\IEEEauthorblockA{
Universidade de São Paulo, Brazil\\
paulormm@ime.usp.br}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
The Linux kernel massive scale (+28 M LoC, +20 K contributors) presents unique 
maintenance challenges. Surprisingly, code duplication remains a persistent issue 
in the kernel codebase, which can hinder its evolution and patching. Academic 
approaches often focus on pairwise comparison of code artifacts, which are not directly applied 
to comprehensive codebase analyses. Other existing free
software\footnote{In this paper, we use the term \textit{Free Software}
interchangeably with \textit{Open Source Software}.} tools explored in 
practice frequently suffer from limited functionality, such as primitive textual 
matching, prove too narrow in scope, or fail to deliver effective results on complex, 
large-scale codebases. Existing solutions generally fail to address the Linux kernel 
specific needs: (1) scalability to handle its size, (2) actionable results for 
developers, and (3) integration with kernel development workflows.
%
This paper presents \textit{ArKanjo}, a novel command-line tool for Linux kernel maintenance 
designed to detect and analyze function-level duplications. Released under the MIT 
license, ArKanjo employs a two-stage architecture consisting of a Preprocessor and 
a Query Responder that separates computationally intensive analysis from efficient 
querying for duplications within large codebases. Pivotal advantages of ArKanjo over 
existing solutions include: (1) optimization for C codebases with kernel-specific 
patterns; (2) preprocessing that enables rapid queries without redundant analysis; 
and (3) prioritization of duplicates that impact maintainability, such as copied 
buggy logic. 
%
We evaluate ArKanjo against real-world duplication cases in recent 
kernel versions, demonstrating its effectiveness in identifying problematic clones 
that generic tools often overlook. By identifying well-defined, manageable 
duplication instances, ArKanjo effectively lowers the barrier for new contributors, 
a capability evidenced by its role in guiding students to make their first code 
improvements to the kernel. ArKanjo offers immediate value to kernel maintainers 
and serves as a replicable model for clone detection in other large-scale free 
software projects.

@@

To fill this gap, this research proposes a new approach embodied in \textit{ArKanjo}, a tool 
designed specifically to identify and facilitate the mitigation of function-level code 
duplications within the Linux kernel. We employ multiple analytical 
methods and ethnographic studies to test it.

@@

\begin{figure}[!t]
\centering
\includegraphics[width=3.5in]{explorer_example}
\caption{Example of the \textit{Duplication Explorer} functionality.}
\label{fig:explorer_ex}
\end{figure}

\begin{figure}[!t]
\centering
\includegraphics[width=3.5in]{function_example}
\caption{Example of the \textit{Function Information} functionality.}
\label{fig:function_ex}
\end{figure}

\begin{figure}[!t]
\centering
\includegraphics[width=3.5in]{relatory_example}
\caption{Example of the \textit{Duplication Report} functionality.}
\label{fig:relatory_ex}
\end{figure}

@@

The Duplication Report functionality provides an overview of
duplicated code within the input codebase. It calculates the number of
duplicated lines per folder in the codebase and presents the information to the user in
a readable format, as demonstrated in Figure~\ref{fig:relatory_ex}.

@@

A discrepancy emerges when detecting non-duplicate pairs between
the two methods. In the BigCloneBench evaluation, we found no false positives (i.e., negative
samples inferred as duplications), whereas the empirical analysis revealed a higher percentage
of false positives. We propose two potential explanations for this discrepancy.
